%\documentclass[a4paper]{article}
\documentclass{llncs}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{multirow}

\usepackage{xstring}
\usepackage{xspace}
\usepackage{amstext}

%\usepackage{showframe}
%\usepackage{times}

\usepackage{listings}
\usepackage{xcolor}
\definecolor{Green}{rgb}{0.1,0.5,0.1}
\lstset{%configuration de listings 
	float=hbp,% 
	language={Python},%
	morekeywords={assert, then},%
	columns=flexible, % 
	keepspaces=true,  % keeps spaces in text, useful for keeping indentation of code
	escapeinside={<@}{@>},%
	tabsize=4, % 
	frame=l, % 
	frameround=tttt, % 
	extendedchars=true, % 
	showspaces=false, % 
	showstringspaces=false, % 
	%numbers=left, % 
	numbersep=5pt,
	breaklines=true, % 
	xleftmargin=15pt, %
	breakautoindent=true, % 
	captionpos=b,% 
    framesep=2pt,
	keywordstyle=\color{blue}, % 
	commentstyle=\color{Green}, % 
} 


% Tikz
\usepackage{tikz}
\usetikzlibrary{arrows,snakes,backgrounds,patterns,matrix,shapes,fit,calc,shadows,plotmarks,intersections,shapes.geometric}

% Include paths
\usepackage{graphicx}
\graphicspath{{assets/}}
\makeatletter
    \def\input@path{{assets/}}
\makeatother

\newcommand{\UNSAFE}{{\color{red!50!black} UNSAFE}\xspace}
\newcommand{\SAFE}{{\color{green!50!black} SAFE}\xspace}
\newcommand{\NA}{{\color{blue!50!black} N/A}\xspace}
\newcommand{\TODO}{{\color{red}\bf TODO}\xspace}
\newcommand{\DiH}{Diffie-Hellman\xspace}
\newcommand{\proverif}{ProVerif\xspace}
\newcommand{\ofmc}{OFMC\xspace}
\newcommand{\spear}{SPEAR II\xspace}
\newcommand{\opcua}{OPC-UA\xspace}
\newcommand{\opc}{OPC\xspace}
\newcommand{\dcom}{DCOM\xspace}
\newcommand{\mms}{MMS\xspace}
\newcommand{\iec}[1]{IEC\xspace#1\xspace}
\newcommand{\iccp}{ICCP\xspace}
\newcommand{\modbus}{MODBUS\xspace}
\newcommand{\profinet}{PROFINET\xspace}
\newcommand{\etherip}{EtherNet/IP\xspace}
\newcommand{\dnp}{DNP3\xspace}
\newcommand{\eg}{{\em e.g.:}\xspace}
\newcommand{\etal}{{\em et al.}\xspace}
\newcommand{\securechan}{{\em OpenSecureChannel}\xspace}
\newcommand{\session}{{\em CreateSession}\xspace}

\title{Short paper: Domain Specific Filtering With Worst-Case Bandwidth}
\author{Maxime Puys \and Jean-Louis Roch \and Marie-Laure Potet}
\institute{Verimag, University Grenoble Alpes,  Gi\`eres, France \\
  \texttt{firstname.lastname@imag.fr}
  \thanks{This work has been partially supported by the LabEx PERSYVAL-Lab
      (ANR-11-LABX-0025) and the project {\em Programme Investissement d’Avenir
      FSN AAP Sécurité Numérique n\textsuperscript{o}3} ARAMIS (P3342-146798).}
}
\date{}

\begin{document}

\renewcommand{\thelstlisting}{\arabic{lstlisting}}

\maketitle

\begin{abstract}
    Industrial systems are publicly the target of cyberattacks since
    Stuxnet \cite{Lan11}.  Nowadays they are increasingly communicating over
    insecure media such as Internet.  Due to their interaction with
    the real world, it is crucial to ensure their security. In this paper, we
    propose a stateful type of domain specific filtering able to keep track
    of the value of predetermined variables. Such filter allows to express rules
    depending on the context of the system. Moreover, it must guarantee bounded
    memory and execution time to be resilient against malicious adversaries.
    Our approach is illustrated on an example.
\end{abstract}

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Stateful Filtering}\label{sec:stateful}
\input{stateful}

\section{Proposed Implementation}\label{sec:impl}

In this Section, we present our filtering model taking the limitations presented
in Section~\ref{sec:stateful} into account.
Then we illustrate it on an example.

\subsection{Description}\label{sec:impl_desc}
\input{impl_desc}

\subsection{Use-Case Example: an Electrical Disconnector}\label{sec:impl_example}
\input{impl_example}

\section{Conclusion}\label{sec:concl}
\input{concl}

\bibliographystyle{unsrt}
\bibliography{phdBiblio}

\end{document}
