
\documentclass[10pt, article]{llncs}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}
%\IEEEoverridecommandlockouts




% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

\usepackage[latin1]{inputenc}
\usepackage{alltt}
\usepackage{listings}
\usepackage{multicol}
\usepackage{verbatim}
\usepackage{url}
\usepackage{xspace}
\usepackage{multicol}
\usepackage{wrapfig}


\newcommand{\shellcmd}[1]{\texttt{\normalsize #1}}


\lstset{  language=C, 
basicstyle=\scriptsize,%\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\scriptsize,%\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1, each line 
                                % will be numbered
numbersep=1pt,                  % how far the line-numbers are from the code
%backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=lines,                   % adds a frame around the code. other options are :  none|leftline|topline|bottomline|lines|single|shadowbox
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
%title=\lstname,                 % show the filename of files included with \lstinputlisting;
                                % also try caption instead of title
framexleftmargin=3mm,           % set the frame 5mm apart from the code and for including line numbers inside the frame.
framexrightmargin=2mm,           % set the frame 1mm apart from the code.
float=t,			% add floating of listings
escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={*,...}            % if you want to add more keywords to the set
          numbers=left,
          captionpos=b,
          caption=\lstname,
}

% The lineno packages adds line numbers. Start line numbering with
% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
% for the whole article with \linenumbers.
%\usepackage{lineno}
\usepackage{graphicx}
\usepackage{graphics}

\usepackage{pdfpages}

%\setlength{\parskip}{3 mm}
%\setlength{\parindent}{3 mm}


\newcommand{\Op}[0]{\textsf{Comp}\xspace}
\newcommand{\Def}[0]{\textsf{Def}\xspace}
\newcommand{\Off}[0]{\textsf{Comp}\xspace}

%\newtheorem{example}{Example}[section]
%\newtheorem{definition}{Definition}[section]



\newcommand{\Refi}{\sqsubseteq}
\newcommand{\combi}{{\sc combines}}
\newcommand{\env}{{\sc env}}      % les variables de la machine

\newcommand{\faults}[1]{{\mbox{Faults}\,(#1)}}



%\title{{Lazart: a symbolic approach to evaluate the impact of fault injections by test inverting,
%in the domain of smart card code.}}

\title{{Lazart: a symbolic approach for evaluating the robustness of secured codes
against  control flow fault injections}}

\author{Marie-Laure Potet, Laurent Mounier, Maxime Puys and Louis Dureuil}

\institute{Laboratoire V\'erimag, Université de Grenoble Alpes \\ pr\'enom.nom@imag.fr}





\begin{document}

\maketitle

% use optional labels to link authors explicitly to addresses:
% \author[label1,label2]{}
% \address[label1]{}
% \address[label2]{}

{\bf Article présenté à ICST 2014\footnote{IEEE International Conference on Software Testing, Verification, and Validation, Cleveland, Ohio, USA, March 31- April 4, 2014}}.
Les cartes à puce sont un vecteur important pour les applications exigeant un haut
niveau de sécurité, par exemple dans le domaine bancaire, de 
l'identité numérique ou du médical. Au vu de la sensibilité des données qu'elles
protègent, le matériel et les applications embarquées doivent être conçues pour résister à des
attaques de haut niveau \cite{CCDB-2009-03-001}. 
%Pour de telles attaques, et en accord vec les Critère Communs \cite{CC06c}, l'analyse de vulnérabilité
%doit couvrir l'état de l'art courant en terme d'attaques.
Nous nous intéressons ici aux attaques par perturbation, consistant en une attaque physique
provoquant une modification du code à exécuter.
%De telles perturbations peuvent par exemple être introduite par une attaque laser.
%Les cartes à puces, et les applications qu'elles embarquent, contiennent donc des contre-mesures
%ayant pour objectif de détecter de telles attaques. 
%Une problèmatique classique pour le développeur est de 
%mettre en place  les bonnes contre-mesures, tout en évitant de trop complexifier le code.
Nous proposons une approche, appelée Lazart,
permettant  d'évaluer la robustesse d'un code contre les injections de fautes visant
des modifications du flot de contrôle.  
%Par rapport aux approches existantes principalement basées
%sur la simulation d'injection de fautes à partir d'une trace d'exécution donnée \cite{BHKL11,Morpho14},  
L'approche Lazart est basée sur une analyse statique du code, dont l'exécution symbolique. On peut ainsi produire
des attaques  mais aussi établir des verdicts d'absence d'attaques. 
De plus le choix d'une analyse statique nous permet de nous intéresser au multi-fautes (plusieurs injections
de faute) tout en maîtrisant la combinatoire que cela introduit.  


L'approche proposée est implémentée dans une chaîne d'outils, basée sur le format LLVM \cite{LLVM}.
Une première étape permet de calculer  les points d'injection de fautes en fonction d'un objectif
d'attaque. 
%Toutes ces injections de faute sont embarquées dans un unique mutant dont l'exécution
Ceux-ci sont embarqués dans un unique mutant dont l'exécution
symbolique  (basée sur le générateur de test concolique Klee  \cite{CadarDE08}), permet 
de calculer les attaques d'ordre inférieur à $n$, pour un $n$ fixé.
Nous présentons les résultats obtenus sur deux implémentations plus ou moins robustes de la vérification
de PIN.
Nous proposons aussi un critère permettant de comparer certaines  attaques. 

%\begin{IEEEkeywords}
%symbolic test generation, fault injection by mutation, smart card vulnerability analysis 
%\end{IEEEkeywords}

% ------------------------------------------------------------------------------------------------



\begin{scriptsize}
\bibliographystyle{alpha}
\bibliography{../ICST14/lazart}
\end{scriptsize}

%\appendix

%\begin{small}
%\input{papersimple.ll}
%\end{small}

%\input{ annexeA.tex}
%\label{annexe}

\end{document}
